---
title: 'EDA for Project 1: Beer and Brewery in the States'
author: "Hao Wang & Alex Salamah"
date: "10/14/2019"
output:
    prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is the code for the EDA deliverable for MSDS Project 1.
# Get the data
```{r, include=FALSE}
setwd("C:/Users/ysalame98/Documents/SMU/Doing DS/MDS-6306-Doing-Data-Science-Fall-2019/Unit 8 and 9 Case Study 1")
BeersDF <- read.csv(file = "Beers.csv", header = TRUE)
BreweriesDF <- read.csv(file = "Breweries.csv", header = TRUE)
AlcoholDF <- read.csv(file = "AlcoholAbuse.csv", header = TRUE, check.names = FALSE)
View(AlcoholDF)
```

# Active library
```{r, include=F}
# Visualization
library(tidyverse)
library(plotly)
library(plyr)
library(dplyr)
library(maps)
library(mapproj)
library(inspectdf)
library(VIM)
library(ggthemes)
library(GGally)

# data wrangling 
library(stringr)
library(tm)
library(knitr)
library(sqldf)

# classification and modeling
library(class)
library(e1071)
library(caret)
```

# 1. How many breweries are present in each state?
```{r}
# Count breweries in each state
table(BreweriesDF$State)

Brew_Num <- BreweriesDF %>%
  group_by(State) %>%
  dplyr::summarize(count=n())
     
# Plot
Brew_Num$State = factor(Brew_Num$State, level = Brew_Num$State[order(-Brew_Num$count)]) 
Brew_Num %>% 
  ggplot(aes(x=State, y=count, fill=State))+geom_bar(stat = "identity", width = 0.8) +
  ylab("Brewery")+ggtitle("Brewery vs State")+
  geom_text(aes(label=count), hjust=0.4, vjust=-0.3, size=3) +
  theme(axis.text.x = element_text(size=7, angle=90))
ggplotly()
```

# 2. Merge beer data with the breweries data. 
# Print the first 6 observations and the last six observations to check the merged file.
# (RMD only, this does not need to be included in the presentation or the deck.)
```{r}
mergeDF <- merge(BeersDF, BreweriesDF, by.x = "Brewery_id", by.y = "Brew_ID")
colnames(mergeDF)[2] = "Beer_name"
colnames(mergeDF)[8] = "Brewery_name"
print("The first 6 observations of the merged file:")
head(mergeDF, n=6)
print("The last 6 observations of the merged file:")
tail(mergeDF, n=6)
```

# 3. Address the missing values in each column.
## Option 1: Simply replacing missing values in IBU/ABV with state median results in significant drop on correlation, from 67% to 49% (Refer to #7)
```{r}
# Check the missing values in the mergeDF
aggr(mergeDF, prop=FALSE, numbers=TRUE)
matrixplot(mergeDF)
marginplot(mergeDF[c("ABV","IBU")], pch = c(20), col = c("darkgray", "red", "blue"))
IBUNADF <-  mergeDF %>% filter(is.na(IBU))
ABVNADF <-  mergeDF %>% filter(is.na(ABV))
BOTHNADF <- mergeDF %>% filter(is.na(ABV) & is.na(IBU))
# Check which state give the most NA's
summary(IBUNADF$State)
summary(mergeDF$State)
summary(ABVNADF$State)
# NADF <- data.frame(State=State, ALL=summary(mergeDF$State), IBUNA=summary(IBUNADF$State), ABVNA=summary(ABVNADF$State))
# NADF$IBUNAperc <- NADF$IBUNA/NADF$ALL
# NADF$ABVNAperc <- NADF$ABVNA/NADF$ALL

# Calculate IBU and ABV median value of each state
IBUmedianState <- mergeDF %>% filter(!is.na(IBU)) %>% group_by(State) %>% dplyr::summarise(IBUmedian=median(IBU))
ABVmedianState <- mergeDF %>% filter(!is.na(ABV)) %>% group_by(State) %>% dplyr::summarise(ABVmedian=median(ABV))

# Replace IBU NA's with state median value
mergeDF_no_IBU_NA <- left_join(mergeDF, IBUmedianState, by = "State")
mergeDF_no_IBU_NA$IBU[is.na(mergeDF_no_IBU_NA$IBU)]  <- as.character(mergeDF_no_IBU_NA$IBUmedian[is.na(mergeDF_no_IBU_NA$IBU)])
aggr(mergeDF_no_IBU_NA, prop=FALSE, numbers=TRUE) # Still 7 observations IBU NA, because there is no any IBU value available for state 

# Replace 7 IBU NA's in SD with grand median value 35.00
mergeDF_no_IBU_NA$IBU = str_replace_na(mergeDF_no_IBU_NA$IBU, replacement =  "35")
mergeDF_no_IBU_NA$IBUmedian = str_replace_na(mergeDF_no_IBU_NA$IBUmedian, replacement =  "35")

# Replace ABV NA's with state median value
mergeDF_noNA <- left_join(mergeDF_no_IBU_NA, ABVmedianState, by = "State")
mergeDF_noNA$ABV[is.na(mergeDF_noNA$ABV)] <- as.character(mergeDF_noNA$ABVmedian[is.na(mergeDF_noNA$ABV)])
aggr(mergeDF_noNA, prop=FALSE, numbers=TRUE)
mergeDF_noNA$IBU <- as.numeric(mergeDF_noNA$IBU)
mergeDF_noNA$ABV <- as.numeric(mergeDF_noNA$ABV)

# Plot
IBU.plotly <- mergeDF_noNA %>% ggplot(aes(x=IBU))+geom_bar()+ggtitle("IBU Distribution") +
  theme(axis.text.x = element_text(size = 6, angle = 90))
ggplotly(IBU.plotly)
ABV.plotly <- mergeDF_noNA %>% ggplot(aes(x=ABV))+geom_bar()+ggtitle("ABV Distribution") +
  theme(axis.text.x = element_text(size = 6, angle = 90))
ggplotly(ABV.plotly)

# Statistic ABV, IBU
summary(mergeDF_noNA)
```

## Option 2: Alternative option is to use Simple Linear Regression model to predict IBU missing values, to keep the correlation between IBU/ABV no change.
```{r}
# Calculate ABV median value of each state
ABVmedianState <- mergeDF %>% filter(!is.na(ABV)) %>% group_by(State) %>%
  dplyr::summarise(ABVmedian=median(ABV))
# Replace 62 NA's in ABV with state level median
mergeDF_no_ABV_NA <- left_join(mergeDF, ABVmedianState, by = "State")

mergeDF_no_ABV_NA$ABV[is.na(mergeDF_no_ABV_NA$ABV)] <-
  mergeDF_no_ABV_NA$ABVmedian[is.na(mergeDF_no_ABV_NA$ABV)]

aggr(mergeDF_no_ABV_NA, prop=FALSE, numbers=TRUE)

# Build simple regression model with 1405 pairs of IBU/ABV to get the correlation coef.
mergeDF_tidy <- mergeDF %>% filter(!is.na(IBU) & !is.na(ABV))
mod <- lm(mergeDF_tidy$IBU ~ mergeDF_tidy$ABV)
summary(mod)
Intercept <- mod$coefficients[1]
Slope <- mod$coefficients[2]

# Add new column predicted_IBU base on correlation coef.
# Change negative predicted_IBU to 0, IBU is measured on a scale from 0 to infinite
mergeDF_no_ABV_NA$predict_IBU <- pmax(Intercept + Slope*mergeDF_no_ABV_NA$ABV, 0)
summary(mergeDF_no_ABV_NA$IBU)

# Replace 1005 NA's in IBU with predicted_IBU
mergeDF_no_ABV_NA$IBU[is.na(mergeDF_no_ABV_NA$IBU)] <-
  mergeDF_no_ABV_NA$predict_IBU[is.na(mergeDF_no_ABV_NA$IBU)]
summary(mergeDF_no_ABV_NA$IBU)

tidyDF <- mergeDF_no_ABV_NA
View(tidyDF)
aggr(tidyDF, prop=FALSE, numbers=TRUE)
```


# 4. Compute the median alcohol content(ABV) and international bitterness unit (IBU) for each state. Plot a bar chart to compare.
```{r}
# plot IBU median distribution using IBUmedianState data frame

## Calculate IBU and ABV median value of each state
IBUmedianState <- tidyDF %>% filter(!is.na(IBU)) %>% group_by(State) %>% dplyr::summarise(IBUmedian=median(IBU))
ABVmedianState <- tidyDF %>% filter(!is.na(ABV)) %>% group_by(State) %>% dplyr::summarise(ABVmedian=median(ABV))
summary(IBUmedianState)
## Plot
IBUmedianState$State <- factor(IBUmedianState$State, level = IBUmedianState$State[order(IBUmedianState$IBUmedian)])
IBUmedianState %>% ggplot(aes(x=State, y=IBUmedian, fill=State)) + geom_bar(stat = "identity") + ggtitle("IBU Median By State")+
  geom_text(aes(label=sprintf("%0.2f", round(IBUmedian,digits = 2))), hjust=-0.5) +
  theme(axis.text.y = element_text(size = 6)) +
  coord_flip()
ggplotly()

# plot ABV median distribution using ABVmedianState data frame
ABVmedianState$State <- factor(ABVmedianState$State, level = ABVmedianState$State[order(ABVmedianState$ABVmedian)])
ABVmedianState %>% ggplot(aes(x=State, y=ABVmedian, fill=State)) + geom_bar(stat = "identity") + ggtitle("ABV Median By State")+
  geom_text(aes(label=sprintf("%0.2f", round(ABVmedian,digits = 2))), hjust=-0.5) +
  theme(axis.text.y = element_text(size = 6)) +
  coord_flip()
ggplotly()
summary(ABVmedianState)

# plot scatter with 51 pairs of median values of ABV and IBU
ABV_IBU_mergeDF <- merge(ABVmedianState, IBUmedianState, by = "State" )
ABV_IBU_mergeDF$State <- factor(ABV_IBU_mergeDF$State, droplevels(ABV_IBU_mergeDF$State))
ABV_IBU_mergeDF %>% ggplot(aes(x=IBUmedian, y=ABVmedian)) + geom_point()
```

# 5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r}
# Calculate IBU and ABV max value of each state
IBUmaxState <- tidyDF %>% filter(!is.na(IBU)) %>% group_by(State) %>% dplyr::summarise(IBUmax=max(IBU))
ABVmaxState <- tidyDF %>% filter(!is.na(ABV)) %>% group_by(State) %>% dplyr::summarise(ABVmax=max(ABV))

# plot IBU max distribution using IBUmedianState data frame
IBUmaxState$State <- factor(IBUmaxState$State, level = IBUmaxState$State[order(IBUmaxState$IBUmax)])
IBUmaxState %>% ggplot(aes(x=State, y=IBUmax, fill=State)) + geom_bar(stat = "identity") + ggtitle("IBU max by State")+
  geom_text(aes(label=sprintf("%0.2f", round(IBUmax,digits = 2))), hjust=+0.5) +
  theme(axis.text.y = element_text(size = 6)) +
  coord_flip()
ggplotly()

# plot ABV max distribution using ABVmaxState data frame
ABVmaxState$State <- factor(ABVmaxState$State, level = ABVmaxState$State[order(ABVmaxState$ABVmax)])
ABVmaxState %>% ggplot(aes(x=State, y=ABVmax, fill=State)) + geom_bar(stat = "identity") + ggtitle("ABV max in States")+
  geom_text(aes(label=sprintf("%0.2f", round(ABVmax,digits = 2))), hjust=+0.5) +
  theme(axis.text.y = element_text(size = 6)) +
  coord_flip()
ggplotly()
```

# 6. Comment on the summary statistics and distribution of the ABV variable.
```{r}
# Create new column for IPA/Ale breakdown
tidyDF$IPA_Ale <- ifelse(str_detect(tidyDF$Style,regex("IPA", ignore_case = TRUE)),"IPA",ifelse(str_detect(tidyDF$Style,regex("Ale",ignore_case = TRUE)),"Ale","Other"))

# Visualization with ABV data
##   Bloxplot
tidyDF %>% ggplot(aes(y=ABV, x=IPA_Ale, fill=as.factor(Ounces))) + geom_boxplot() +
  xlab("Ale Types (IPA or ALE)") + ylab("ABV") + ggtitle("Boxplot of ABV vs Ale Type")
tidyDF %>% ggplot(aes(y=ABV, x=as.factor(Ounces), fill=IPA_Ale)) + geom_boxplot() +
  xlab("Ounces") + ylab("ABV") + ggtitle("Boxplot of ABV vs Ounces")
##   Histogram
tidyDF %>% ggplot(aes(x=ABV)) + geom_histogram(fill="blue", color="black")
##   Barchart
tidyDF %>% ggplot(aes(x=ABV, fill=IPA_Ale)) + geom_bar(color="black") + xlim(c(0.025,0.100)) +
  xlab("Alcohol By Volume (ABV)") + ylab("Beer Name Count") + ggtitle("Ale/IPA Weight vs ABV")
tidyDF %>% ggplot(aes(x=ABV, fill=IPA_Ale)) + geom_bar(color="black", position = "fill") + xlim(c(0.025,0.100)) +
  xlab("Alcohol By Volume (ABV)") + ylab("Beer Name Count%") + ggtitle("Ale/IPA Weight% vs ABV")
tidyDF %>% ggplot(aes(x=ABV, fill=as.factor(Ounces))) + geom_bar(color="black") + xlim(c(0.025,0.100)) +
  xlab("Alcohol By Volume (ABV)") + ylab("Beer Name Count") + ggtitle("Ounce Weight vs ABV")

# Visualization with IBU data
##   Bloxplot
tidyDF %>% ggplot(aes(y=IBU, x=IPA_Ale, fill=as.factor(Ounces))) + geom_boxplot() +
  xlab("Ale Types (IPA or ALE)") + ylab("IBU") + ggtitle("Boxplot of IBU vs Ale Type")
tidyDF %>% ggplot(aes(y=IBU, x=as.factor(Ounces), fill=IPA_Ale)) + geom_boxplot() +
  xlab("Ounces") + ylab("IBU") + ggtitle("Boxplot of IBU vs Ounces")
##   Barchart
tidyDF %>% ggplot(aes(x=IBU, fill=IPA_Ale)) + geom_bar(color="black") +
  xlab("International Bitterness Units (IBU)") + ylab("Beer Name Count") + ggtitle("Ale/IPA Weight vs IBU")
tidyDF %>% ggplot(aes(x=IBU, fill=IPA_Ale)) + geom_bar(color="black", position = "fill") +
  xlab("International Bitterness Units (IBU)") + ylab("Beer Name Count%") + ggtitle("Ale/IPA Weight% vs IBU")
tidyDF %>% ggplot(aes(x=IBU, fill=as.factor(Ounces))) + geom_bar(color="black") +
  xlab("International Bitterness Units (IBU)") + ylab("Beer Name Count") + ggtitle("Ounce Weight vs IBU")
# Scatter plot ABV vs. IBU
tidyDF %>% ggplot(aes(x=IBU, y=ABV)) + geom_point(shape=0) + ggtitle("ABV vs IBU") + geom_smooth(method = lm)
tidyDF %>% ggplot(aes(x=IBU, y=ABV, color=IPA_Ale)) + geom_point() + ggtitle("ABV vs IBU") + geom_smooth(method = lm)
tidyDF %>% ggplot(aes(x=IBU, y=ABV, color=IPA_Ale)) + geom_point() + ggtitle("ABV vs IBU") + facet_wrap(~State)
tidyDF %>% ggplot(aes(x=IBU, y=ABV, color=IPA_Ale)) + geom_point() + ggtitle("ABV vs IBU") + facet_wrap(~as.factor(Ounces))
```

# 7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.
```{r}
# Scatter plot and Pearson's correlation with original data
mergeDF %>% filter(!is.na(IBU) & !is.na(ABV)) %>%
  ggplot(aes(x=ABV, y=IBU)) + 
  geom_point(aes(color=State), size=1, position="jitter") + 
  geom_smooth(method = lm) + 
  theme_wsj() +
  xlab("ABV") + ylab("IBU") +
  ggtitle("Relationship between IBU and ABV")
ggplotly()
mergeDF_tidy <- mergeDF %>% filter(!is.na(IBU) & !is.na(ABV))
cor.test(mergeDF_tidy$IBU, mergeDF_tidy$ABV)
mod <- lm(mergeDF_tidy$IBU ~mergeDF_tidy$ABV)
summary(mod)

# Scatter plot and Pearson's correlation with tidy data (no missing value)
tidyDF %>% ggplot(aes(x=ABV, y=IBU)) + 
  geom_point(aes(color=State), size=1, position="jitter") + 
  geom_smooth(method = lm) + 
  ggtitle("Relationship between IBU and ABV")
ggplotly()
cor.test(tidyDF$IBU, tidyDF$ABV)
mod1 <- lm(tidyDF$IBU ~ tidyDF$ABV)
summary(mod1)

#Correlations
Corl <- inspect_cor(mergeDF)
show_plot (Corl) + ggtitle("Correlation Between All Attributes")
# Sctter Plot
mergeDF %>% ggplot(aes(x=IBU, y=ABV)) + geom_point(aes(color=State), size=1, position="jitter") + geom_smooth()

# Plot GGpairs
tidyDF %>%
select(ABV, IBU, IPA_Ale) %>%
ggpairs(mapping = aes(color = IPA_Ale))
```

# 8. Budweiser would also like to investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA).  You decide to use KNN clustering to investigate this relationship.  Provide statistical evidence one way or the other. You can of course assume your audience is comfortable with percentages … KNN is very easy to understand.
```{r}
table(tidyDF$IPA_Ale) # Ale:963, IPA:571, Other:876
# Check the Ale: 2nd time with regex(): # Ale:1007, IPA:571, Other:832

tidyDF.knn <- tidyDF %>% filter(IPA_Ale == "IPA" | IPA_Ale == "Ale")

## 
set.seed(1234)
splitPerc = .75
iterations = 50
numks = 30

masterAcc = matrix(nrow = iterations, ncol = numks)
  
for(j in 1:iterations)
{
  accs = data.frame(accuracy = numeric(30), k = numeric(30))
  trainIndices = sample(1:dim(tidyDF.knn)[1],round(splitPerc * dim(tidyDF.knn)[1]))
  train = tidyDF.knn[trainIndices,]
  test = tidyDF.knn[-trainIndices,]
  for(i in 1:numks)
    {
    classifications = knn(train[,c(4,5)],test[,c(4,5)],train$IPA_Ale, prob = TRUE, k = i)
    table(classifications,test$IPA_Ale)
    CM = confusionMatrix(table(classifications,test$IPA_Ale))
    masterAcc[j,i] = CM$overall[1]
  }
}

MeanAcc = colMeans(masterAcc)
maxMeanAcc = max(MeanAcc) # k=?
plot(seq(1,numks,1),MeanAcc, type = "l")

classifications5 = knn(train[,c(4,5)],test[,c(4,5)],train$IPA_Ale, prob = TRUE, k = 5)
table(classifications5,test$IPA_Ale)
CM5 = confusionMatrix(table(classifications5,test$IPA_Ale))
CM5
```

```{r}
set.seed(1234)
splitPerc = .80
iterations = 50
numks = 30

masterAcc = matrix(nrow = iterations, ncol = numks)
  
for(j in 1:iterations)
{
  accs = data.frame(accuracy = numeric(30), k = numeric(30))
  trainIndices = sample(1:dim(tidyDF.knn)[1],round(splitPerc * dim(tidyDF.knn)[1]))
  train = tidyDF.knn[trainIndices,]
  test = tidyDF.knn[-trainIndices,]
  for(i in 1:numks)
    {
    classifications = knn(train[,c(4,5)],test[,c(4,5)],train$IPA_Ale, prob = TRUE, k = i)
    table(classifications,test$IPA_Ale)
    CM = confusionMatrix(table(classifications,test$IPA_Ale))
    masterAcc[j,i] = CM$overall[1]
  }
}

MeanAcc = colMeans(masterAcc)
maxMeanAcc = max(MeanAcc) # k=?
plot(seq(1,numks,1),MeanAcc, type = "l")

classifications5 = knn(train[,c(4,5)],test[,c(4,5)],train$IPA_Ale, prob = TRUE, k = 6)
table(classifications5,test$IPA_Ale)
CM5 = confusionMatrix(table(classifications5,test$IPA_Ale))
CM5
```

# Check Budweiser profile
## Neither Budweiser nor Anheuser-Busch can be found in list of Brewery_name
## We cannot build Budweiser profile with the available data set
## Benchmarking with competitors will not be realistic
```{r}
tidyDF$BudweiserOrOther <- ifelse(str_detect(tidyDF$Brewery_name,regex("Budweiser", ignore_case = TRUE)),"Budweiser","Other")
table(tidyDF$BudweiserOrOther)
head(tidyDF %>% filter(BudweiserOrOther=="Budweiser"),n=50)
```


# 9. Knock their socks off!  Find one other useful inference from the data that you feel Budweiser may be able to find value in.  You must convince them why it is important and back up your conviction with appropriate statistical evidence.
```{r}
  lookup = data.frame(abb=state.abb, state= state.name )
  #View(lookup)
  # add new column to tidyDF with lower case state codes to user to merge with lookup DF
  tidyDF$abb <-   tidyDF$State
  #View(tidyDF)
  
  # create a new data set with BrewriesDF and lookup data
  BrewriesDF1 = merge(tidyDF,lookup,"abb")
  #View(BrewriesDF1)
  # count # of sate in Acu2 data frame
  BrewriesMPD= count(BrewriesDF1,"state")
  #View(BrewriesMPD)
  # Change the name of the count column to Brewreies Count
  colnames(BrewriesMPD)[2] = "BrewriesCnt"
  #change the names of the states to lower case.  
  BrewriesMPD$region <- tolower(BrewriesMPD$state)
  # get the longi/lat of the states in order to draw the map
  states <- map_data("state")
  #View(states)
  #merge the states long/lat data with AcuMap data
  map.df <- left_join(states,BrewriesMPD, by = "region")
  #View(map.df)
  # Plot the data on map
  ggplot(map.df, aes(long, lat, group = group))+
    geom_polygon(aes(fill = BrewriesCnt ), stat= "identity", color = "white")+
    #geom_text(aes(label=state), hjust=0.5, vjust=0.5, size=3)  
    #geom_text(aes(label = BrewriesMPD$count, )) +
    scale_fill_gradientn (colours= (heat.colors(3)),na.value="grey90")+
     ggtitle( "Breweries by State", subtitle = "Geopositional Analysis")

# Analyze External Datasource AlcoholDF
  AlcoholDF$region <- tolower(AlcoholDF$State)
  AlcoholDF$NumTot <- as.numeric(AlcoholDF$Total)
  AlcoholDF$chaTot<- as.character(AlcoholDF$NumTot)
  AlcoholMap.df <- left_join(states,AlcoholDF, by = "region")
  View(AlcoholMap.df)
  ggplot(AlcoholMap.df, aes(long, lat, group = group)) +
    geom_polygon(aes(fill = GreaterThan12Est), stat= "identity", color = "white") +
    #geom_text(aes(label=state), hjust=0.5, vjust=0.5, size=3)  
    #geom_text(aes(label = BrewriesMPD$count, )) +
    scale_fill_gradientn (colours= (heat.colors(3)),na.value="grey90")+
     ggtitle( "Alcohol Abuse by State", subtitle = "Geopositional Analysis")

```


